\documentclass[
	%parspace, % Add vertical space between paragraphs
	%noindent, % No indentation of first lines in each paragraph
	%nohyp, % No hyphenation of words
	%twoside, % Double sided format
	%draft, % Quicker draft compilation without rendering images
	%final, % Set final to hide todos
]{elteikthesis}[2024/04/26]

% The minted package is also supported for source highlighting
% See elteikthesis_minted.tex for example
%\usepackage[newfloat]{minted}

% Toggle: set \UPMtrue for UPM submission, \UPMfalse for ELTE
\newif\ifUPM
\UPMtrue % change to \UPMtrue for UPM university submission

% Document's metadata
\title{Lane Boundary Extraction from High-Density LiDAR Point Clouds for HD Mapping} % title
\date{2026} % year of defense

% Author's metadata
\author{Haála Kada Iván}
\degree{Data Science MSc}

% Supervisor(s)' metadata
\ifUPM
	\supervisor{Alberto Mozo, Full Professor} % internal supervisor's name and title
	\affiliation{Mathematical Modeling and Biocomputing Group, Universidad Politécnica de Madrid} % internal supervisor's affiliation
\else
	\supervisor{Hajder Levente, Habil. Associate Professor} % internal supervisor's name and title
	\affiliation{Institute of Computer Science, Department of Algorithms and Their Applications, Eötvös Loránd University} % internal supervisor's affiliation
\fi
\extsupervisor{José Daniel Padrón, PhD} % external supervisor's name
\extaffiliation{Data Science, Maps Advanced Analytics, TomTom N.V.} % external supervisor's affiliation

% University's metadata
\ifUPM
	\university{Universidad Politécnica de Madrid} % university's name
	\faculty{Escuela Técnica Superior de Ingenieros Informáticos} % faculty's name
	\department{} % department's name (required; leave empty if not applicable)
	\city{Madrid} % city
	\logo{upm_logo} % logo
	\logotwo{etsiinf_logo} % second logo
\else
	\university{Eötvös Loránd University} % university's name
	\faculty{Faculty of Informatics} % faculty's name
	\department{} % department's name (required; leave empty if not applicable)
	\city{Budapest} % city
	\logo{elte_cimer_szines} % logo
\fi

% Add bibliography file
\addbibresource{elteikthesis.bib}

% The document
\begin{document}

% Set document language
%\documentlang{hungarian}
\documentlang{english}

% List of todos (not in the final document)
%\listoftodos[\todolabel]

% Title page (mandatory)
\maketitle
% Topic declaration page (mandatory) - can also be attached instead
%\includepdf{topicdeclaration.pdf}

% Abstract
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
This thesis presents a geometry-based offline processing pipeline for extracting vectorized lane boundary geometries from high-density mobile LiDAR point clouds for High Definition (HD) mapping applications.
The proposed framework addresses the challenge of generating accurate lane-level ground truth data for HD map validation without relying on rasterized representations or large-scale annotated training datasets.
Since the provided LiDAR data lacked vehicle trajectory information, four trajectory reconstruction algorithms were developed and evaluated, including a novel Pulse Line Intersection method exploiting the LiDAR packet structure.
The pipeline further combines spatial binning, ground segmentation, intensity-based lane marking extraction, road surface detection, and sliding-window curve fitting to reconstruct continuous lane geometries directly from raw point clouds.
Experimental evaluation on approximately 10\,km of highway LiDAR data demonstrates that the proposed methods can reliably extract both solid and dashed lane boundaries while preserving centimetre-level geometric fidelity.
The resulting vectorized lane representations enable scalable and precise cross-section-based HD map validation workflows.
\cleardoublepage

% Table of contents (mandatory)
\tableofcontents
\cleardoublepage

% Main content
\input{chapters/1_introduction.tex}
\cleardoublepage

\input{chapters/2_literature_rev.tex}
\cleardoublepage

\input{chapters/3_methodology.tex}
\cleardoublepage

\input{chapters/4_evaluation.tex}
\cleardoublepage

\input{chapters/5_future_work.tex}
\cleardoublepage

% Acknowledgements (optional) - in case your thesis received funding or would like to express special thanks to someone
\chapter*{\acklabel}
\addcontentsline{toc}{chapter}{\acklabel}

I would like to express my gratitude to TomTom N.V.~for the opportunity to work on this interesting topic within their Maps Advanced Analytics division.
Their provision of hardware infrastructure, LiDAR datasets, and a stimulating research environment were instrumental in making this thesis possible.

\ifUPM\else
I am also grateful to my internal supervisor, Hajder Levente, for his patient guidance, insightful feedback, and support throughout the research and writing process.

My friend Dávid deserves special thanks for his encouragement during the most challenging periods of this work, providing both moral support and perspective when needed.

\fi
Most importantly, I wish to extend my heartfelt appreciation to both my Mother and Father.
Their constant support, encouragement, and belief in me throughout all my life and studies have been the foundation upon which this achievement stands.
This thesis is, in many ways, a reflection of their dedication and love.


% Appendices (optional) - useful for detailed information in long tables, many and/or large figures, etc.
\appendix
% \input{chapters/appendix.tex}
% \cleardoublepage

% Bibliography (mandatory)
\phantomsection
\addcontentsline{toc}{chapter}{\biblabel}
\printbibliography[title=\biblabel]
\cleardoublepage

% List of figures (optional) - useful over 3-5 figures
% \phantomsection
% \addcontentsline{toc}{chapter}{\lstfigurelabel}
% \listoffigures
% \cleardoublepage

% List of tables (optional) - useful over 3-5 tables
% \phantomsection
% \addcontentsline{toc}{chapter}{\lsttablelabel}
% \listoftables
% \cleardoublepage

% List of algorithms (optional) - useful over 3-5 algorithms
% \phantomsection
% \addcontentsline{toc}{chapter}{\lstalgorithmlabel}
% \listofalgorithms
% \cleardoublepage

% List of codes (optional) - useful over 3-5 code samples
% \phantomsection
% \addcontentsline{toc}{chapter}{\lstcodelabel}
% \lstlistoflistings
% \cleardoublepage

% List of symbols (optional)
%\printnomenclature

\end{document}
